---
title: Go Help Yourself?
subtitle: Teacher Educator Attributions of Failure and the Provision of Social Support
title-block-banner: "#3459e6"
author:
  - name: David Simon
    orcid: 0009-0005-6451-3790
    email: da.simon@dipf.de
    affiliations:
      - name: DIPF | Leibniz Institute for Research and Information in Education
date: today
date-format: DD MMMM YYYY
editor: source
execute:
  warning: false
  message: false
  cache: true
format: 
  html:
    echo: true
    theme: zephyr
    mainfont: Arial
    fontsize: 1.2em
    code-block-bg: true
    code-block-border-left: "#3459e6"
    toc: true
    toc-location: left
    toc-depth: 4
    embed-resources: true
    code-fold: true
    code-tools: true
    code-link: true
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
bibliography: references.bib
---

```{css}
#| echo: false
p {
  text-align: justify
}
```

## Information

```{r}
# Input: b_sab_mzp2_2024_R_data_file.csv
#
# Output: --
#
# Contents: (1) Read Data
#           (2) Measures
#           (3) Results

####  ------------- Load Packages ------------- ####
library(here)           # find files
library(tidyverse)      # for data management
library(psych)          # for scale construction, EFA, CFA
library(lavaan)         # for SEM
```

## Data Analysis

### 1 Read Data

Data:

```{r}
data <- read.csv(here("01_data/b_sab_mzp2_2024_R_data_file.csv"))
```

### 2 Measures

#### 2.1 Exploratory Factor Analysis (EFA): Social Support Scales  
The items measuring social support were taken from Richter at al. [-@richter2011soziale] and Schulz and Schwarzer [-@BSS]. We adapted to the educators' perspective. Items were added to each facet according to our theoretical understanding of social support. An EFA was calculated to determine which items represented the three facets of social support best.

##### Dataevaluation

Data for EFA:

```{r}
data_efa <- data %>%
  dplyr::select(sozunt_ins5:sozunt_ins2)
```

*Kaiser–Meyer–Olkin (KMO) Test:*

- 0.00 - 0.49: Unacceptable
- 0.50 - 0.59: Miserable
- 0.60 - 0.69: Mediocre
- 0.70 - 0.79: Middling
- 0.80 - 0.89: Meritorious
- 0.90 - 1.00: Marvelous

```{r}
psych::KMO(data_efa)
```

Deletion of item `sozunt_ins4`, who failed the KMO test (< 0.50):

```{r}
data_efa <- data_efa %>%
  dplyr::select(-sozunt_ins4)
```

*Bartlett Test of Sphericity:*

```{r}
psych::cortest.bartlett(data_efa)$p.value
```

##### Exploratory Factor Analysis

*Step 1: Plausibility of a three factor solution*

Get Eigenvalues:

```{r}
eig <- eigen(cor(data_efa, use = "complete.obs")) # get eigenvalues
eig$values
```
Show scree plot:

```{r}
psych::scree(data_efa, pc = FALSE)  # plot scree plot 
```

The eigenvalue method (“Kaiser’s rule”) suggested four factors. The scree plot suggested two factors. In line with our theoretical understanding of the construct we started with three factors first, assuming correlated factors: 

*Step 2: Factor extraction*

```{r}
efa <- psych::fa(data_efa, nfactors = 3, rotate = "oblimin")
loadings <- efa$loadings
loadings
```

*Step 3: Item removal*

Only variables with loadings of .32 and above were interpreted as suggested as a first rule of thumb by Tabachnick and Fidell [-@tabachnick2001using]. 

Consequently `sozunt_emo1`, `sozunt_ins3` and `sozunt_ins2` were not further considered:

```{r}
data_efa <- data_efa %>%
  dplyr::select(-sozunt_emo1, -sozunt_ins3, -sozunt_ins2)
```

The EFA was repeated:

```{r}
efa <- psych::fa(data_efa, nfactors = 3, rotate = "oblimin")
loadings <- efa$loadings
loadings
```

*Step 4: Interpretation of factors*  

Factors were interpreted given the intended theoretical assignment of items to factors and a final cut of .45 regarding the loadings, which equals a 20% variance overlap between variable and factor and was characterised as "fair" by Comray (-@Comrey1962) and examplified in Tabachnick and Fidell [-@tabachnick2001using]. 

This resulted in the following three factors, which correspond to the three facets of social support:

- Factor 1: Informational support ("Tips and advice") was best represented through the items `sozunt_inf2`, `sozunt_inf3`, `sozunt_inf4` and  `sozunt_inf5`,

- Factor 2: Emotional support ("Comfort and nearness") was best represented through the items `sozunt_emo2`, `sozunt_emo3` and `sozunt_emo4`.

- Factor 3: Instrumental support ("Goods and materials") was best represented through the items `sozunt_ins1` and `sozunt_ins5`,

Adjustment of data accordingly:

```{r}
data_efa <- data_efa %>%
  dplyr::select(sozunt_inf2, sozunt_inf3, sozunt_inf4, sozunt_inf5,
                sozunt_emo2, sozunt_emo3, sozunt_emo4,
                sozunt_ins1, sozunt_ins5)
```

*Step 5: Final solution*

```{r}
efa <- psych::fa(data_efa, nfactors = 3, rotate = "oblimin")
loadings <- efa$loadings
loadings
psych::fa.diagram(loadings)
```
`sozunt_inf1` does not meet the cutoff criterion anymore. Since the item has no cross-loading we kept it and would only discard it, if reliability could be improved by dropping it, which was not the case: 

*Step 6: Evaluate Reliability*

```{r}
inf_sup_alpha <- data %>% 
  dplyr::select(sozunt_inf2, sozunt_inf3, sozunt_inf4, sozunt_inf5) %>%
  psych::alpha(title = "sozunt_inf")

emo_sup_alpha <- data %>%  
  dplyr::select(sozunt_emo2, sozunt_emo3, sozunt_emo4) %>%
  psych::alpha(title = "sozunt_emo")

ins_sup_alpha <- data %>% 
  dplyr::select(sozunt_ins1, sozunt_ins5) %>%
  psych::alpha(title = "sozunt_ins")
```

- Alpha informational support: `r inf_sup_alpha$total$std.alpha`
- Alpha emotional support: `r emo_sup_alpha$total$std.alpha`
- Alpha instrumental support: `r ins_sup_alpha$total$std.alpha`  

*Step 7: Does this solution fit the data?*

Conducting a maximum likelihood factor analysis showed the solution fits the data well:

```{r}
three_efa_ml <- psych::fa(data_efa, nfactors = 3, rotate = "oblimin", fm = "ml")

three_efa_ml_chisq <- three_efa_ml$STATISTIC  # chi-square statistic
three_efa_ml_pval <- three_efa_ml$PVAL  # p-value
```

- Chi-square: `r three_efa_ml_chisq`
- P-value: `r three_efa_ml_pval`

And further parsimoniousness would not:

```{r}
two_efa_ml <- fa(data_efa, nfactors = 2, rotate = "oblimin", fm = "ml")

two_efa_ml_chisq <- two_efa_ml$STATISTIC  # chi-square statistic
two_efa_ml_pval <- two_efa_ml$PVAL  # p-value
```

- Chi-square: `r two_efa_ml_chisq`
- P-value: `r two_efa_ml_pvalchiquadrat statistik

#### 2.2 Confirmatory Factor Analysis (CFA):  Self-Efficacy Scale
Tickell and Klassen [-@tickell2024developing] developed a  scale for measuring teacher mentoring self-efficacy beliefs and suggest a two-factor model, with factors named as *Pedagogical Practices* and *Professional Relationships*. We tested this factor structure in our German translation of the questionnaire:

##### Confirmatory Factor Analysis

Data for CFA:

```{r}
data_cfa <- data %>%
  dplyr::select(seffsab_1:seffsab_9)
```

*Step 1: Fitting a two-factor model*

```{r}
# specify the model
cfa_model <- ' pae =~ seffsab_1 + seffsab_2 + seffsab_20 + seffsab_14 + 
                      seffsab_6 + seffsab_28 + seffsab_48 + seffsab_39 +
                      seffsab_46 + seffsab_47 + seffsab_4
               rel =~ seffsab_29 + seffsab_44 + seffsab_27 + seffsab_42 + 
                      seffsab_9 '

# fit the model
fit_cfa <- cfa(cfa_model, data = data_cfa)
```

Model fit was poor:

```{r}
# CFI cut-off: .95; SRMR cut-off: .09
fitMeasures(fit_cfa, c("chisq", "df", "pvalue", "cfi", "srmr"))
```

*Step 2: Evaluate Reliability*

In order to improve the model-fit we evaluated reliability:

```{r}
sef_pea_alpha <- data %>% 
  dplyr::select(seffsab_1, seffsab_2, seffsab_20, seffsab_14, seffsab_6, 
                seffsab_28, seffsab_48, seffsab_39, seffsab_46, seffsab_47,
                seffsab_4) %>%
  psych::alpha(title = "sef_pea_alpha")

sef_rel_alpha <- data %>%  
  dplyr::select(seffsab_29, seffsab_44, seffsab_27, seffsab_42, seffsab_9) %>%
  psych::alpha(title = "sef_rel_alpha")
```

We were interested whether reliability improves if an item is dropped and if that might as well improve model fit. Alpha could not be improved in case of *Professional Relationships* nor *Relationships*: 

```{r}
sef_pea_alpha$alpha.drop
sef_rel_alpha$alpha.drop
```

But in case of *Professional Relationships* several items could be dropped without worsening. As we believed the translation of `seffsab_48` might be not suitable, we discarded this item and reran the CFA:

```{r}
data_cfa <- data_cfa %>%
  dplyr::select(-seffsab_48)

# specify the model
cfa_model_2 <- ' pae =~ seffsab_1 + seffsab_2 + seffsab_20 + seffsab_14 + 
                        seffsab_6 + seffsab_28 + seffsab_39 + seffsab_46 + 
                        seffsab_47 + seffsab_4
                 rel =~ seffsab_29 + seffsab_44 + seffsab_27 + seffsab_42 + 
                        seffsab_9 '

# fit the model
fit_cfa_2 <- cfa(cfa_model_2, data = data_cfa)
fitMeasures(fit_cfa_2, c("chisq", "df", "pvalue", "cfi", "srmr"))
```

Model fit improved, but was still not good enough;

```{r}
# CFI cut-off: .95; SRMR cut-off: .09
fitMeasures(fit_cfa_2, c("chisq", "df", "pvalue", "cfi", "srmr"))
```

Looking at the reliability output again, we saw no room for further improvement:

```{r}
sef_pea_alpha <- data %>% 
  dplyr::select(seffsab_1, seffsab_2, seffsab_20, seffsab_14, seffsab_6, 
                seffsab_28, seffsab_39, seffsab_46, seffsab_47, seffsab_4) %>%
  psych::alpha(title = "sef_pea_alpha")

sef_pea_alpha$alpha.drop
```

Although the model fit was not as good as desired, we considered it to be sufficient.

#### 2.3 Compute Scale Scores

```{r}
# Factor vignette, dummy for each vignette
data <- data %>%
  dplyr::mutate(vig = case_when((random == 1) ~ 1,
                                (random == 2) ~ 2,
                                (random == 3) ~ 3,
                                TRUE ~ as.numeric(NA)),
                vig1 = ifelse(vig == 1, 1, 0),
                vig2 = ifelse(vig == 2, 1, 0),
                vig3 = ifelse(vig == 3, 1, 0))


data$vig <- as.factor(data$vig)

## Attribution 

data <- data %>%
  dplyr::mutate(int = int_1,  # internality
                sta = sta_1,  # stability 
                glo = glo_1,  # globality
                con = kon_1)  # controllability

# Attribution style: Int, Sta, Glo
data <- data %>% 
  dplyr::mutate(isg = rowMeans(subset(data, select = c(int, sta, glo)), na.rm = TRUE))

isg_alpha <- data %>% 
  dplyr::select(int, sta, glo) %>%
  psych::alpha(title = "isg_alpha")


## Affect 

# Negative affect, positive affect
data <- data %>% 
  dplyr::mutate(pos_aff = rowMeans(subset(data, select = c(affekt_1, affekt_3, 
                                                           affekt_4, affekt_6, 
                                                           affekt_10, affekt_11,
                                                           affekt_12, affekt_15)),
                                   na.rm = TRUE),
                neg_aff = rowMeans(subset(data, select = c(affekt_2, affekt_5, 
                                                           affekt_7, affekt_8, 
                                                           affekt_13, affekt_14,
                                                           affekt_16)),
                                   na.rm = TRUE),
                ang = affekt_5,   # Anger
                gui = affekt_7,   # Guilt
                sha = affekt_14)  # Shame


## Social Support

# Informational support: "Hinweise und Ratschläge"
data <- data %>% 
  dplyr::mutate(inf_sup = rowMeans(subset(data, select = c(sozunt_inf2,
                                                           sozunt_inf3,
                                                           sozunt_inf4, 
                                                           sozunt_inf5)),
                                   na.rm = TRUE),

# Emotional support: "Trost und Nähe"
                emo_sup = rowMeans(subset(data, select = c(sozunt_emo2, 
                                                           sozunt_emo3, 
                                                           sozunt_emo4)),
                                   na.rm = TRUE),
                
## Instrumental support: "Güter und Materialien"                
                ins_sup = rowMeans(subset(data, select = c(sozunt_ins1, 
                                                           sozunt_ins5)),
                                 na.rm = TRUE))

## Self Efficacy

# Self Efficacy, Factor 1: Pedagogical Practices   
data <- data %>% 
  dplyr::mutate(seff_f1 = rowMeans(subset(data, select = c(seffsab_1, seffsab_2, 
                                                           seffsab_20, seffsab_14, 
                                                           seffsab_6, seffsab_28, 
                                                           seffsab_39, seffsab_46, 
                                                           seffsab_47, seffsab_4)),
                                   na.rm = TRUE),

# Self Efficacy, Factor 2: Relationships
                seff_f2 = rowMeans(subset(data, select = c(seffsab_29,seffsab_44,
                                                           seffsab_27,seffsab_42, 
                                                           seffsab_9)),
                                   na.rm = TRUE))
```

#### 2.4 Glossary

*Attributions*   
adapted from [@zurattributionsstil]

- `int`: Internality
- `sta`: Stability
- `glo`: Globality
- `con`: Controllability

*Affect*  
after [@breyer2016deutsche]

- `pos_aff`: Positive Affect
- `neg_aff`: Negative Affect
- `ang`: Anger
- `gui`: Guilt
- `sha`: Shame

*Social Support*  
adapted from [@richter2011soziale]

- `inf_sup`: Informational support
- `ins_sup`: Instrumental support
- `emo_sup`: Emotional Support

*Self Efficacy*  
adapted from [@tickell2024developing]

- `seff_f1`: Pedagogical Practices
- `seff_f2`: Relationships

### 3 Results

#### 3.1 Attributional Theory

Weiner (1985) identified three causal dimensions.  

One dimension of causes is their *locus*, or location within or outside of the actor (locus of control). Ability and effort are similar in locus, both being internal to the actor. In addition, they differ from causes such as chance or task difficulty, which are considered to be located in the environment, external to the actor.  

A second dimension of causes is *stability*, which depicts a cause as
constant or changing over time. Some causes, including ability, are
viewed as stable over time. On the other hand, chance or luck fluctuates
over time and is therefore unstable. Effort also is typically regarded as
unstable, although labeling oneself or others as lazy or industrious
implies stability in effort expenditure. 

The third known causal dimension is *controllability*, or whether a cause is subject to volitional influence. Attribution theorists have proposed that ability, although internal to the person, is not subject to volitional control and change, whereas effort expenditure is considered amenable to personal control – “it could be otherwise.”

Zitiert nach [@graham2020attributional]

The *locus of control* dimension is particularly relevant to emotional reactions. Internal attributions for undesirable events or behaviors are frequently associated with self-focused negative emotions, such as guilt and shame. External attributions for the same behaviors and outcomes are generally associated with externally focused negative emotions, such as anger and resentment.

Zitiert nach [@weiner1985attributional]

Was würde das für die SAB bedeuten: Liegt die Ursache in der LAA, dann liegt es nicht an ihnen und sie reagieren ärgerlich - "die LAA sollen gefälligst liefern". (?)

Ähnliche Effekte gibt es in Bezug auf "controllability" wie folgt:

In her earlier paper, Graham (1984) claimed that affect is the mediator between perceptions of *controllability* and subsequent helping behavior. As she explained, perceived uncontrollability gives rise to sympathy, which in turn promotes the commitment to help. In contrast, “perceived controllability elicits anger, and anger then leads to the withholding of help” Graham (1984)

Graham (1990) suggested that failing students gain information about the causes of their failure from the affective displays of teachers. This attributional information may sometimes be quite direct, but at other times it may be “subtly, indirectly, and even unknowingly conveyed” (p. 18). Referring to the findings of a series of studies, the same author claimed that communication of pity and the offering of unsolicited help to the student might function as low-ability or even incompetence cues. On the other hand, when failure is attributed to insufficient effort, the attributional message communicated to the student by the teacher’s anger may have positive consequences.

A study among Israeli primary school teachers (Butler, 1994) supported Graham’s findings. In this study it was found that teachers were more likely to respond to the low-effort students with anger and demands for better work and to the low-ability students with pity and offers of help.

Zitiert nach [@georgiou2002teacher]

- H1a: Controllability predicts witholding of help
- H1b: Internality predicts witholding of help
- H1c: The relationships are mediated by anger

(?)

- Problem "Internality" könnte low-effort und low-ability sein, was aber gegensätzliche Reaktionen auslösen würde. Wie könnte das trotzdem abgebildet werden?
- Andere Emotionen außer Ärger finden bisher keine Beachtung

```{r}
model1 <- '# direct effect
           inf_sup ~ int + con
           ins_sup ~ int + con
           emo_sup ~ int + con 
           # mediators
           ang ~ int + con
           inf_sup ~ ang
           ins_sup ~ ang
           emo_sup ~ ang'

fit1 <- lavaan::sem(model1, data = data, missing = "fiml")

summary(fit1)
```

#### 3.2 Self Efficacy

Mentors with high levels of efficacy might believe that they could control teacher students achievement and motivation. Thus, mentors with high sense of efficacy are more willing to take responsibility for failures and are more likely to assist low achieving teacher students during failure situations.

- H2: Mentors with high level of efficacy perceive the situation as more controllable (Controllability)
- H3: Relationship Self-Efficacy positively predicts the provision of social support

(?)

- H2b: Mentors with high level of efficacy perceive the situation as externally caused (Internality = Locus of Control)
- vielleicht: H2c: Mentors with high level of efficacy perceive the situation as unstable (Stability)

```{r}
model2 <- 'inf_sup ~ seff_f2 + seff_f1
           ins_sup ~ seff_f2 + seff_f1
           emo_sup ~ seff_f2 + seff_f1

           con ~ seff_f2 + seff_f1'

fit2 <- lavaan::sem(model2, data = data, missing = "fiml")

summary(fit2)
```

#### 3.3 Both

```{r}
model3 <- '# direct effect
           inf_sup ~ int + con + seff_f2 + seff_f1
           ins_sup ~ int + con + seff_f2 + seff_f1
           emo_sup ~ int + con + seff_f2 + seff_f1
           # mediators
           ang ~ int + con
           inf_sup ~ ang
           ins_sup ~ ang
           emo_sup ~ ang
           # else
           con ~ seff_f2 + seff_f1'

fit3 <- lavaan::sem(model3, data = data, missing = "fiml")

summary(fit3)

# CFI cut-off: .95; SRMR cut-off: .09
fitMeasures(fit3, c("chisq", "df", "pvalue", "cfi","srmr"))
```

*Might there be a relationship between emotions and self-efficacy?*

```{r}
model4 <- 'ang ~ seff_f2 + seff_f1
           gui ~ seff_f2 + seff_f1
           sha ~ seff_f2 + seff_f1'

fit4 <- lavaan::sem(model4, data = data, missing = "fiml")
summary(fit4)
```

#### 3.4 Descriptives: Vignettes

*Were the situations perceived as a failure?*

"How much do you see the lessons you have experienced as a failure or a success?"

- 1, 2, 3: Failure
- 4, 5, 6: Success

```{r}
data_vig1 <- data %>% 
  dplyr::filter(vig == 1) 

data_vig2 <- data %>% 
  dplyr::filter(vig == 2) 

data_vig3 <- data %>% 
  dplyr::filter(vig == 3) 

tab1 <- table(data_vig1$check_1)
tab2 <- table(data_vig2$check_1)
tab3 <- table(data_vig3$check_1)

barplot(tab1, main = "Vignette 1")
barplot(tab2, main = "Vignette 2")
barplot(tab3, main = "Vignette 3")
```

*Were the situations cause equally attributed?*

```{r}
tab1 <- data.frame(int = mean(data_vig1$int, na.rm = T),
                   sta = mean(data_vig1$sta, na.rm = T),
                   glo = mean(data_vig1$glo, na.rm = T),
                   con = mean(data_vig1$con, na.rm = T))

tab2 <- data.frame(int = mean(data_vig2$int, na.rm = T),
                   sta = mean(data_vig2$sta, na.rm = T),
                   glo = mean(data_vig2$glo, na.rm = T),
                   con = mean(data_vig2$con, na.rm = T))

tab3 <- data.frame(int = mean(data_vig3$int, na.rm = T),
                   sta = mean(data_vig3$sta, na.rm = T),
                   glo = mean(data_vig3$glo, na.rm = T),
                   con = mean(data_vig3$con, na.rm = T))

barplot(as.matrix(tab1), main = "Vignette 1")
barplot(as.matrix(tab2), main = "Vignette 2")
barplot(as.matrix(tab3), main = "Vignette 3")
```

## Next Steps

- Nur mit Vignette 1 und 3 rechnen (?)
- Positive Emotionen anschauen
- Hätte es eine CFA sein sollen? Ich habe eine EFA gerechnet, weil ich mir ja Items ausgedacht habe und wissen wollte, welche davon die passenden sind.

## References

- Correlation Matrix

::: {#refs}
:::
