---
title: "Go Help Yourself:"
subtitle: "Teacher Educator Attributions of Failure and the Witholding of Social Support" 
title-block-banner: true
author: DS
date: today
date-format: "DD MMMM YYYY"
editor: source
execute:
  echo: true
  warning: false
  message: false
  cache: true
format: 
  html:
    echo: true
    theme: flatly
    fontsize: 1.1em
    toc: true
    toc-location: left
    toc-depth: 3
    embed-resources: true
    code-fold: false
    code-tools: true
    code-link: true
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
bibliography: references.bib
---

## Information

```{r}
# Input: b_sab_mzp2_2024_R_data_file.csv
#
# Output: --
#
# Contents: (1) Read Data
#           (2) Measures
#           (3) Results

####  ------------- Load Packages ------------- ####
library(here)           # find files
library(tidyverse)      # for data management
library(psych)          # for scale construction
library(tidySEM)        # for SEM
library(lavaan) 

```

## Data Analysis

### (1) Read Data

```{r}
data <- read.csv(here("01_data/b_sab_mzp2_2024_R_data_file.csv"))
```

### (2) Measures

#### (2a) Exploratory Factor Analysis: Social Support Scale
The items measuring social support were taken from [@richter2011soziale] and adapted to the educators' perspective. Items were added to each question according to the theoretical understanding. An EFA was calculated to determine which items represented the three facets of social support best.  

##### Dataevaluation

*Evaluation of missingness:*

```{r}
data_efa <- data %>%
  dplyr::select(sozunt_ins5:sozunt_ins2)

desc <- tidySEM::descriptives(data_efa)
desc <- desc[, c("name", "type", "n", "missing", "unique", "mode")]
desc
```

Mean substitution of missing data:

```{r}
mean_substitution <- function(df){
  n_cols <- ncol(df)
  
  for(i in 1:n_cols){
  # get the current column
  column_i <- df[, i]
  
  # get the mean of the current column
  mean_i <- mean(column_i, na.rm = TRUE)
  
  # get the NAs in the current column
  NAs_i <- which(is.na(column_i))
  
  # report the number of NAs
  N_NAs <- length(NAs_i)
  
  # replace the NAs in the current column
  column_i[NAs_i] <- mean_i
  
  # replace the original column with the
  ## updated columns
  df[, i] <- column_i
  
  }
  
  return(df)
}

data_efa <- mean_substitution(data_efa)
```

*Kaiser–Meyer–Olkin (KMO) Test:*

- 0.00 - 0.49: Unacceptable
- 0.50 - 0.59: Miserable
- 0.60 - 0.69: Mediocre
- 0.70 - 0.79: Middling
- 0.80 - 0.89: Meritorious
- 0.90 - 1.00: Marvelous

```{r}
psych::KMO(data_efa)
```

Deletion of those items, who failed the KMO test (< 0.50):

```{r}
data_efa <- data_efa %>%
  dplyr::select(-sozunt_ins4)
```

*Bartlett Test of Sphericity:*

```{r}
psych::cortest.bartlett(data_efa)$p.value
```

##### Exploratory Factor Analysis

*Step 1: How many factors to extract?*

```{r}
psych::scree(data_efa, pc = FALSE)       # plot scree plot 
psych::fa.parallel(data_efa, fa = "both")  # check against parallel analysis, missing data is treated pairwise
```

The scree plot suggests two factors. Parallel analysis is revealing up to five factors.

*Step 2: Extract factors*

```{r}
fit <- stats::factanal(data_efa, 3, rotation = "varimax")
print(fit, digits = 2, cutoff = 0.4, sort = TRUE)
```


*Step 3: Plot factors*
```{r}
loadings <- fit$loadings
psych::fa.diagram(loadings)
```
According to this

- Informational support ("Hinweise und Ratschläge") is best represented through the items `sozunt_inf3`, `sozunt_inf4` and  `sozunt_inf5`  

- Instrumental support ("Güter und Materialien") is best represented through the items `sozunt_ins5` and `sozunt_ins1`

- Emotional support ("Trost und Nähe") is best represented through the items `sozunt_emo2`, `sozunt_emo3` and  `sozunt_emo4`

*Step 4: Evaluate Reliability*
```{r}
inf_sup_alpha <- data %>% 
  dplyr::select(sozunt_inf3, sozunt_inf4, sozunt_inf5) %>%
  psych::alpha(title = "sozunt_inf")

ins_sup_alpha <- data %>% 
  dplyr::select(sozunt_ins5, sozunt_ins1) %>%
  psych::alpha(title = "sozunt_ins")

emo_sup_alpha <- data %>%  
  dplyr::select(sozunt_emo2, sozunt_emo3, sozunt_emo4) %>%
  psych::alpha(title = "sozunt_emo")

inf_sup_alpha$total
ins_sup_alpha$total
emo_sup_alpha$total
```

Notiz:
Für "Instrumental support" ist alpha .56. - Richter et al. schreiben "Auf die Erfassung der instrumentellen Unterstützung von Seiten des Mentors wurde verzichtet, da die einschlägige Literatur zum Mentoring diese Form der Unterstützung bisher kaum thematisiert (diese Einschränkung wird im Diskussionsteil angesprochen)."

Eine Konsequenz könnte sein, die Skala nicht weiter zu berücksichtigen. Womöglich spielt diese Art der Unterstützung tatsächlich keine Rolle.

#### (2b) Compute Scale Scores

```{r}
# Factor vignette, dummy for each vignette
data <- data %>%
  dplyr::mutate(vig = case_when((random == 1) ~ 1,
                                (random == 2) ~ 2,
                                (random == 3) ~ 3,
                                TRUE ~ as.numeric(NA)),
                vig1 = ifelse(vig == 1, 1, 0),
                vig2 = ifelse(vig == 2, 1, 0),
                vig3 = ifelse(vig == 3, 1, 0))


data$vig <- as.factor(data$vig)

### Attribution 

data$int <- data$int_1 # internality
data$sta <- data$sta_1 # stability
data$glo <- data$glo_1 # globality

data$con <- data$kon_1 # controllability


# Attribution style: Int, Sta, Glo
data <- data %>% 
  dplyr::mutate(isg = rowMeans(subset(data, select = c(int, sta, glo)), na.rm = TRUE))

isg_alpha <- data %>% 
  dplyr::select(int, sta, glo) %>%
  psych::alpha(title = "isg_alpha")

### Affect 

# Negative affect, positive affect
data <- data %>% 
  dplyr::mutate(pos_aff = rowMeans(subset(data, select = c(affekt_1, affekt_3, 
                                                           affekt_4, affekt_6, 
                                                           affekt_10, affekt_11,
                                                           affekt_12, affekt_15)),
                                   na.rm = TRUE),
                neg_aff = rowMeans(subset(data, select = c(affekt_2, affekt_5, 
                                                           affekt_7, affekt_8, 
                                                           affekt_13, affekt_14,
                                                           affekt_16)),
                                   na.rm = TRUE))

data$ang <- data$affekt_5   # Anger
data$gui <- data$affekt_7   # Guilt
data$sha <- data$affekt_14  # Shame

### Social Support

## Informational support: "Hinweise und Ratschläge"
data <- data %>% 
  dplyr::mutate(inf_sup = rowMeans(subset(data, select = c(sozunt_inf3, 
                                                           sozunt_inf4, 
                                                           sozunt_inf5)),
                                   na.rm = TRUE),
                
## Instrumental support: "Güter und Materialien"                
                ins_sup = rowMeans(subset(data, select = c(sozunt_ins5, 
                                                           sozunt_ins1)),
                                 na.rm = TRUE),

## Emotional support: "Trost und Nähe"
                emo_sup = rowMeans(subset(data, select = c(sozunt_emo2, 
                                                           sozunt_emo3, 
                                                           sozunt_emo4)),
                                   na.rm = TRUE))

### Self Efficacy

data <- data %>% 
  dplyr::mutate(seff = rowMeans(subset(data, select = c(seffsab_1, seffsab_2, 
                                                        seffsab_20, seffsab_14, 
                                                        seffsab_6, seffsab_28,
                                                        seffsab_48, seffsab_39,
                                                        seffsab_46, seffsab_47,
                                                        seffsab_4, seffsab_29,
                                                        seffsab_44, seffsab_27,
                                                        seffsab_42, seffsab_9)),
                                na.rm = TRUE))

# Self Efficacy, Factor 1: Pedagogical Practices
data <- data %>% 
  dplyr::mutate(seff_f1 = rowMeans(subset(data, select = c(seffsab_1, seffsab_2, 
                                                        seffsab_20, seffsab_14, 
                                                        seffsab_6, seffsab_28,
                                                        seffsab_48, seffsab_39,
                                                        seffsab_46, seffsab_47,
                                                        seffsab_4)),
                                na.rm = TRUE))

# Self Efficacy, Factor 2: Relationships
data <- data %>% 
  dplyr::mutate(seff_f2 = rowMeans(subset(data, select = c(seffsab_29,seffsab_44, 
                                                        seffsab_27,seffsab_42, 
                                                        seffsab_9)),
                                na.rm = TRUE))
```

#### (2c) Glossary

*Attributions* 
adapted from [@zurattributionsstil]

- `int`: Internality
- `sta`: Stability
- `glo`: Globality
- `con`: Controllability
- `isg`: Attribution style: internal, stable and global

*Affect* 
after [@breyer2016deutsche]

- `pos_aff`: Positive Affect
- `neg_aff`: Negative Affect
- `ang`: Anger
- `gui`: Guilt
- `sha`: Shame

*Social Support*
adapted from [@richter2011soziale]

- `inf_sup`: Informational support
- `ins_sup`: Instrumental support
- `emo_sup`: Emotional Support

*Self Efficacy*
adapted from [@tickell2024developing]

- `seff_f1`: Pedagogical Practices
- `seff_f2`: Relationships

### (3) Results

#### (3a) Attributional Theory



## References

::: {#refs}
:::
